---
title: "Laboration 6 Writing Fast R Code"
author: "Linus Kage and Malke Marouki Group 4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Laboration_6_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Importing the library and creating knapsack object:

```{r setup, warning=FALSE, message=FALSE}
library(lab6knapsack)
library(Rcpp)

RNGversion(min(as.character(getRversion()),"3.5.3"))

##old sampler used for backward compatibility
## suppressWarnings() can be used so that the above warning is not displayed

set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

```

## Methods:

### Brute force search:

Question: How much time does it takes to run the algorithm for n = 16 objects?

```{r}
microbenchmark::microbenchmark(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500),times=10)
```

When n = 16, it takes about 0.3 to 0.5 seconds to run the algorithm.

### Dynamic programming

Question: How much time does it takes to run the algorithm for n = 500 objects?


```{r}
microbenchmark::microbenchmark(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500),times=3)
```

Using 500 objects and the dynamic programming algorithm, it takes about 25-40 seconds to run the algorithm.

### Greedy heuristic

Question: How much time does it takes to run the algorithm for n = 1000000 objects?

```{r}
RNGversion(min(as.character(getRversion()),"3.5.3"))

##old sampler used for backward compatibility
## suppressWarnings() can be used so that the above warning is not displayed

set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects_1 <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

microbenchmark::microbenchmark(greedy_knapsack(x = knapsack_objects_1[1:1000000,], W = 3500),times=3)

```

Using the greedy heuristics and 1000000 objects, the time it takes to run the algorithm is about 0.3 seconds.


---

## Profile your code and optimize your code

We try to use `ifelse()`, which is vectorized, to try to speed up the computation.

```{r}
knapsack_dynamic_opt <- function(x, W){
  
  n <- nrow(x)
  
  m <- matrix(data = 0, nrow = n, ncol = W)
  
  for(i in 2:n){
    
    for(j in 1:W){
      
      #Vectorized ifelse statement
      m[i,j] <- ifelse(x[i,1] > j, m[i-1,j], max(m[i-1,j], m[i-1,j-x$w[i]] + x$v[i]))
    }
  }
  
  
  
  n <- nrow(x)
  c <- W
  element <- vector()
  while(c > 0){
    if(m[n,c] != m[n-1,c]){
      element <- c(element, n)
      c <- c - x$w[n]
      
      
      
    }
    n <- n-1
    if(n == 1) break
  }
  
  return(list(value = round(max(m)), element = rev(element)))
}

```

```{r ,eval=TRUE}
microbenchmark::microbenchmark(knapsack_dynamic_opt(x = knapsack_objects_1[1:500,], W = 3500),times=3)
```

Question: What performance gain could you get by trying to improving your code?

When running `profvis()`, we noticed a bottleneck in the dynamic programming. We used a nested for-loop and if-else statement which took a long time to compute. The results show that using the `ifelse()` makes the algorithm run slower compared to the previous code that did not use vectorized `ifelse()`.

## (*) Implentation in Rcpp

```{r,eval=FALSE}
#Run with C++ code.
microbenchmark::microbenchmark(knapsack_dynamic(x = knapsack_objects[1:16,], W = 3500,fast = TRUE),times=3)

```

Question: What performance gain could you get by using Rcpp and C++?

Using C++ with the nested for-loops, the run time is much lower compared to the code without C++ implementation. The run time of the algorithm without C++ code is about 25-30 seconds, and with C++ code it takes about 0,3 seconds which is a big improvement.





